---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# persuasio

<!-- badges: start -->
[![CRAN status](https://www.r-pkg.org/badges/version/persuasio)](https://CRAN.R-project.org/package=persuasio)
<!-- badges: end -->

`persuasio` estimates and bounds persuasion effects in instrumental variable
settings with binary outcomes. You provide the outcome, the treatment, and
the instrument, tell `persuasio` which estimand you want (average or local
persuasion rate), and it takes care of the bounds and inference. Based on
Jun and Lee (2023) <https://doi.org/10.1086/724114>.

## Installation

You can install the development version of persuasio from [GitHub](https://github.com/) with:

``` r
# install.packages("pak")
pak::pak("xinrui33/persuasio")
```

## Quick Example

```{r example}
library(persuasio)
## basic example code

# Average persuasion rate (APR): normal inference
persuasio(
  est     = "apr",
  varlist = c("voteddem_all", "readsome", "post"),
  data    = GKB,
  level   = 0.80,
  method  = "normal"
)

# Local persuasion rate (LPR): bootstrap inference
persuasio(
  est     = "lpr",
  varlist = c("voteddem_all", "readsome", "post"),
  data    = GKB,
  level   = 0.80,
  method  = "bootstrap",
  nboot   = 1000
)
```

## Learn more

See `vignette("persuasio")` for a full walkthrough including covariates,
model specifications, and the relationship between estimands.
