---
title: "Simple Causal Inference of Persuasion Effects with 'persuasio'"
author: "Xinrui Chen"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{persuasio}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## What is this for?

This shows the basic usage of the \pkg{persuasio} package for the causal inference of average persuasion effects (APR) and local persuasion effects (LPR) following Jun and Lee (2023).

The package estimates persuasion effects in settings with
a two-stage causal structure:

$$Z_i \text{ (instrument)} \rightarrow T_i \text{ (exposure)} \rightarrow Y_i \text{ (outcome)}$$

where $Z_i$ is a binary instrument (e.g. random assignment, policy eligibility),
$T_i$ is actual exposure to a persuasive message (e.g. media consumption,
campaign contact), and $Y_i$ is a binary behavioural outcome (e.g. voting,
donating, purchasing). For an overview of the empirical settings motivating
these estimands, see DellaVigna and Gentzkow (2010, *Annual Review of
Economics*,
[doi:10.1146/annurev.economics.102308.124309](https://doi.org/10.1146/annurev.economics.102308.124309)).

### When is \pkg{persuasio} applicable?

The framework applies whenever the empirical setting features:
- **A binary or discretized outcome** — vote/not vote, donate/not donate
- **A persuasive channel** — media exposure, information shocks, campaigns,
policy eligibility
- **A credible instrument** — Randomized control trial (RCT) assignment, quasi-random rollout
- **Imperfect compliance** — not everyone assigned receives treatment, or some
untreated still receive exposure (fuzzy design)

Estimation requires at minimum $P(Y \mid Z)$ and $P(T \mid Z)$; individual-level
data on $(Y, T, Z)$ is the ideal case. `persuasio` also accommodates partial
settings, where $Y, Z$ come from one source and $T, Z$ from another are also
accommodated via the `"yz"` estimator.

### Key estimands
**Average Persuasion Rate (APR)** The APR is the population-wide persuasion rate and is the analoguous of the average treatment effect (ATE); it is generally partially identified and
reported as sharp bounds.
| **Local Persuasion Rate (LPR)** The LPR conditions on the complier
subpopulation, those for whom $Z_i$ changes $T_i$, and is analogous of the local
average treatment effect (LATE) in the instrument variable literature.

## \pkg{persuasio} TL;DR

`persuasio` is the recommended function for all estimation and inference.
The lower-level functions (`aprlb`, `aprub`, `lpr4ytz`, `calc4persuasio`)
are available for direct access when only a specific component is needed.

### Choosing an estimator (`est`) in `persuasio`

| Situation                                              | `est` argument          |
|--------------------------------------------------------|-------------------------|
| APR bounds + CI, have treatment variable               | `est = "apr"`           |
| LPR point estimate + CI                                | `est = "lpr"`           |
| APR bounds + CI, no treatment variable                 | `est = "yz"`            |
| Bounds from summary statistics                         | `est = "calc"`          |

### Choosing an inference method in `persuasio`

| Situation                                              | Recommended `method`    |
|--------------------------------------------------------|-------------------------|
| No covariates, large sample                            | `"normal"`              |
| Covariates present                                     | `"bootstrap"`           |
| Small sample regardless of covariates                  | `"bootstrap"`           |
| Need reproducible bootstrap results                    | `"bootstrap"` + `seed`  |

### Choosing a model in `persuasio` when covariates are present

| Identifying assumption                                          | `model`             |
|-----------------------------------------------------------------|---------------------|
| Homogeneous covariate effects across instrument groups          | `"no_interaction"`  |
| Heterogeneous covariate effects across instrument groups        | `"interaction"`     |
| Unsure                                                 | `"no_interaction"` (default) |

### Package setup

```{r setup}
library(persuasio)
```

We load the data set included in the \pkg{persuasio} package to demonstrate the functions.

```{r, error = TRUE}
load("data/GKB.rda")
```

## `persuasio`: Unified Wrapper for Persuasion Effect Estimation

`persuasio` is the recommended entry point for the \pkg{persuasio} package.
Rather than calling `aprlb`, `aprub`, `lpr4ytz`, or `calc4persuasio`
directly, `persuasio` provides a single consistent interface that handles
input parsing and dispatches to the correct estimator based on the `est`
argument.

### Argument overview

| Argument  | Purpose                                              | Default         |
|-----------|------------------------------------------------------|-----------------|
| `est`     | Which estimand to target (`"apr"`, `"lpr"`, `"yz"`, `"calc"`) | — |
| `varlist` | Variable names in order: `y`, `t`, `z`, then optional `x` | — |
| `data`    | Data frame                                           | —               |
| `level`   | Confidence level                                     | `0.95`          |
| `model`   | `"no_interaction"` or `"interaction"`                | `"no_interaction"` |
| `method`  | `"normal"` or `"bootstrap"`                          | `"normal"`      |
| `nboot`   | Bootstrap replications (only when `method = "bootstrap"`) | `1000`     |

### `varlist` ordering

Variables must always be supplied in the order `c(y, t, z, x, ...)`.
For `est = "yz"` or `est = "calc"`, the treatment variable `t` is not
required and should be omitted.

### Examples using the GKB dataset

**Example 1: Average persuasion rate (APR)**

```{r, error = TRUE}
persuasio(
  est     = "apr",
  varlist = c("voteddem_all", "readsome", "post"),
  data    = GKB,
  level   = 0.80,
  method  = "normal"
)
```

Dispatches to `aprlb` and `aprub` to produce lower and upper bounds on
the APR with asymptotic normal confidence intervals at the 80% level.

**Example 2: Local persuasion rate (LPR)**

```{r, error = TRUE}
persuasio(
  est     = "lpr",
  varlist = c("voteddem_all", "readsome", "post"),
  data    = GKB,
  level   = 0.80,
  method  = "normal"
)
```

Dispatches to `lpr4ytz` to estimate the persuasion effect among compliers and
returns standard errors and confidence intervals assuming asymptotic normality.

**Example 3: Outcome-instrument bounds with covariate and bootstrap**

```{r, error = TRUE}
persuasio(
  est     = "yz",
  varlist = c("voteddem_all", "readsome", "post", "MZwave2"),
  data    = GKB,
  level   = 0.80,
  model   = "interaction",
  method  = "bootstrap",
  nboot   = 1000
)
```

The `"yz"` estimand uses only the outcome and instrument (no treatment
variable), making it suitable when treatment data are unavailable. `MZwave2` is
passed as a covariate with `model = "interaction"`, allowing heterogeneous
covariate effects across instrument groups. Bootstrap inference with 1,000
replications preferable to the asymptotic normal approximation.


## Base estimators

This section details the base-level functions of the \pkg{persuasio} package.

| Situation                                              | Base function         |
|--------------------------------------------------------|-----------------------|
| Lower bound on APR only                                | `aprlb`               |
| Upper bound on APR only                                | `aprub`               |
| Local persuasion rate (LPR) for compliers              | `lpr4ytz`             |
| Bounds from published group means, no microdata        | `calc4persuasio`      |


### `calc4persuasio`: Bounds from Summary Statistics

`calc4persuasio` computes the lower and upper bounds on the average persuasion
rate (APR) directly from treatment group means. This function is useful when only
published summary statistics are available, or as a quick sanity check
against`aprlb` and `aprub`.

The function requires:

| Argument | Meaning                                          |
|----------|--------------------------------------------------|
| `y1`     | Mean of binary outcome `y` when instrument `z = 1` |
| `y0`     | Mean of binary outcome `y` when instrument `z = 0` |
| `e1`     | Mean of binary treatment `t` when instrument `z = 1` |
| `e0`     | Mean of binary treatment `t` when instrument `z = 0` |

#### Examples using the GKB dataset

Since `calc4persuasio` works on summary statistics, we first compute the group means manually:

```{r, error = TRUE}
voteddem_all_0 <- mean(GKB$voteddem_all[GKB$post == 0], na.rm = TRUE)
voteddem_all_1 <- mean(GKB$voteddem_all[GKB$post == 1], na.rm = TRUE)

readsome_0 <- mean(GKB$readsome[GKB$post == 0], na.rm = TRUE)
readsome_1 <- mean(GKB$readsome[GKB$post == 1], na.rm = TRUE)
```

Then pass them to `calc4persuasio`:

```{r, error = TRUE}
calc4persuasio(
  y1 = voteddem_all_1,
  y0 = voteddem_all_0,
  e1 = readsome_1,
  e0 = readsome_0
)
```

### `aprlb`: Lower Bound on the Average Persuasion Rate

The `aprlb` function estimates the **lower bound** on the average persuasion
rate (APR) following Jun and Lee (2023).

The function requires:

| Argument | Role                        | Must be binary? |
|----------|-----------------------------|-----------------|
| `y`      | Outcome                     | Yes             |
| `z`      | Instrument                  | Yes             |
| `x`      | Covariates (optional)       | No              |

#### Examples using the GKB dataset
**Example 1: No covariates**

```{r, error = TRUE}
aprlb(data = GKB, y = "voteddem_all", z = "post")
```

This estimates the lower bound on the average persuasion rate without any
covariates, returning a coefficient, standard error, and 95% CI.

**Example 2: With a covariate**

```{r, error = TRUE}
aprlb(data = GKB, y = "voteddem_all", z = "post", x = "MZwave2")
```

With covariates, the function fits two separate models on the `z =
1` and `z = 0` subgroups. Standard errors are not available in this case.

**Example 3: By subgroup**

```{r, error = TRUE}
gkb_groups <- split(GKB, GKB$MZwave2)
lapply(gkb_groups, function(sub_data) {
  aprlb(data = sub_data, y = "voteddem_all", z = "post")
})
```

This example computes `aprlb` separately for each value of `MZwave2` without
covariates, producing estimates with standard errors for each subgroup.

### `aprub`: Upper Bound on the Average Persuasion Rate

The `aprub` function estimates the **upper bound** on the average persuasion
rate (APR) following Jun and Lee (2023).

The function requires:

| Argument | Role                        | Must be binary? |
|----------|-----------------------------|-----------------|
| `y`      | Outcome                     | Yes             |
| `t`      | Treatment                   | Yes             |
| `z`      | Instrument                  | Yes             |
| `x`      | Covariates (optional)       | No              |


#### Examples using the GKB dataset
**Example 1: No covariates**

```{r, error = TRUE}
aprub(data = GKB, y = "voteddem_all", t = "readsome", z = "post")
```

Without covariates, `aprub` estimates the upper bound directly from the
data and returns a coefficient, standard error, and 95% confidence interval.

**Example 2: With a covariate**

```{r, error = TRUE}
aprub(data = GKB, y = "voteddem_all", t = "readsome", z = "post", x = "MZwave2")
```

With covariates, the function fits separate models on the `z = 1` and `z = 0`
subgroups. Standard errors are not available in this case.

**Example 3: By subgroup**

```{r, error = TRUE}
gkb_groups <- split(GKB, GKB$MZwave2)
lapply(gkb_groups, function(sub_data) {
  aprub(data = sub_data, y = "voteddem_all", t = "readsome", z = "post")
})
```

This example computes the no-covariate estimations of `aprub` within each level
of `MZwave2`, returning a full estimate with standard errors for each subgroup.

### `lpr4ytz`: Local Persuasion Rate

`lpr4ytz` estimates the **local persuasion rate (LPR)** — the average
persuasion effect among compliers. While `aprlb` and `aprub` bound the average
persuasion rate (APR) across the full population, the LPR focuses on the compliers.

The function requires:

| Argument | Role                        | Must be binary? |
|----------|-----------------------------|-----------------|
| `y`      | Outcome                     | Yes             |
| `t`      | Treatment                   | Yes             |
| `z`      | Instrument                  | Yes             |
| `x`      | Covariates (optional)       | No              |

#### Examples using the GKB dataset
**Example 1: No covariates**

```{r, error = TRUE}
lpr4ytz(
  data = GKB,
  y    = "voteddem_all",
  t    = "readsome",
  z    = "post"
)
```

Without covariates, `lpr4ytz` returns the LPR estimate along with a standard
error and 95% confidence interval.

**Example 2: With a covariate, interaction model**

```{r, error = TRUE}
lpr4ytz(
  data = GKB,
  y    = "voteddem_all",
  t    = "readsome",
  z    = "post",
  x    = "MZwave2"
)
```

With covariates, the `model = "interaction"` fits separate models on the
`z = 1` and `z = 0` suvgroups, allowing heterogeneous covariate effects across
instrument groups. Standard errors are not available in this case.

**Example 3: With a covariate, no-interaction model**

```{r, error = TRUE}
lpr4ytz(
  data  = GKB,
  y     = "voteddem_all",
  t     = "readsome",
  z     = "post",
  x     = "MZwave2",
  model = "no_interaction"
)
```

The default `model = "no_interaction"` fits a single model. This is a more
parsimonious specification and is appropriate when there is no strong prior
reason to expect heterogeneous covariate effects.

## Inference wrappers

| Situation                                              |  Function               |
|--------------------------------------------------------|-------------------------|
| APR bounds + CI, with treatment variables              | `persuasio4ytz`         |
| APR bounds + CI, no treatment variables                | `persuasio4yz`          |
| LPR point estimate + CI                                | `persuasio4ytz2lpr`     |

Each inference wrapper calls the corresponding base function(S) for point
estimation, then adds confidence intervals via the chosen inference `method`:

- `persuasio4ytz` calls `aprlb` and `aprub` jointly, combining the two
  bounds into a single interval using either the Stoye (2009) normal
  correction or bootstrap resampling.
- `persuasio4yz` follows the same structure as `persuasio4ytz` but omits
  the treatment variable, using only the reduced-form relationship between
  outcome and instrument.
- `persuasio4ytz2lpr` calls `lpr4ytz` for the LPR point estimate and
  wraps it with a standard normal or bootstrap confidence interval.

In all three cases, `persuasio` is the recommended entry point and
dispatches to the appropriate wrapper internally. Call the wrappers
directly only when finer control over arguments is needed.

### `persuasio4ytz`: Confidence Intervals for the Average Persuasion Rate

When `est = "apr"` in `persuasio`, `persuasio4ytz` is working under the hood. It
combines the lower and upper bound estimates from `aprlb` and `aprub` into a
single confidence interval, using either an asymptotic normal approximation or
bootstrap resampling.

#### Examples using the GKB dataset
**Example 1: No covariates, asymptotic normal approximation**

```{r, error = TRUE}
persuasio4ytz(
  data   = GKB,
  y      = "voteddem_all",
  t      = "readsome",
  z      = "post",
  method = "normal",
  level  = 0.80
)
```

The asymptotic normal method is appropriate here since no covariates are present
and analytic standard errors are available from both `aprlb` and `aprub`.

**Example 2: No covariates, bootstrap inference**

```{r, error = TRUE}
persuasio4ytz(
  data   = GKB,
  y      = "voteddem_all",
  t      = "readsome",
  z      = "post",
  method = "bootstrap",
  level  = 0.80,
  nboot  = 1000
)
```

Bootstrap inference with 1,000 replications provides an alternative to the
normal approximation. Results should be close to Example 1 in large samples.
Note that the default `nboot = 50` is suitable only for quick checks. Use larger
`nboot` for more precise inference.

**Example 3: With covariate, interaction model, bootstrap inference**

```{r, error = TRUE}
persuasio4ytz(
  data   = GKB,
  y      = "voteddem_all",
  t      = "readsome",
  z      = "post",
  x      = "MZwave2",
  model  = "interaction",
  method = "bootstrap",
  level  = 0.80,
  nboot  = 1000
)
```

When covariates are present, analytic standard errors are unavailable, so
`method = "bootstrap"` is required. The `model = "interaction"` specification
allows heterogeneous covariate effects of `MZwave2` across the `z = 1` and `z =
0` subgroups.

#### Relationship to `persuasio`

```{r eval=FALSE}
# These two calls are equivalent
persuasio(
  est     = "apr",
  varlist = c("voteddem_all", "readsome", "post"),
  data    = GKB,
  level   = 0.80,
  method  = "normal"
)

persuasio4ytz(
  data   = GKB,
  y      = "voteddem_all",
  t      = "readsome",
  z      = "post",
  method = "normal",
  level  = 0.80
)
```

### `persuasio4ytz2lpr`: Confidence Intervals for the Local Persuasion Rate

When `est = "lpr"` in `persuasio`, `persuasio4ytz2lpr` is working under the
hood. It is the inference companion to `lpr4ytz`, wrapping the LPR point
estimator and adding confidence intervals via either a delta-method normal
approximation or bootstrap resampling.

When covariates are present, `lpr4ytz` returns `se = NA` and `method =
"bootstrap"` must be used. Set `seed` for reproducible bootstrap results.

#### Examples using the GKB dataset
**Example 1: No covariates, normal inference**

```{r, error = TRUE}
persuasio4ytz2lpr(
  data   = GKB,
  y      = "voteddem_all",
  t      = "readsome",
  z      = "post",
  method = "normal",
  level  = 0.80
)
```

The function uses the delta-method standard error from `lpr4ytz` to construct an
80% confidence interval under the normal approximation. This is the fastest
option and appropriate for large samples without covariates.

**Example 2: No covariates, bootstrap inference**

```{r, error = TRUE}
persuasio4ytz2lpr(
  data   = GKB,
  y      = "voteddem_all",
  t      = "readsome",
  z      = "post",
  method = "bootstrap",
  level  = 0.80,
  nboot  = 1000
)
```

Bootstrap inference with 1,000 replications provides an alternative to the
normal approximation. Results should align closely with Example 1 in large
samples. The default `nboot = 50` is only appropriate for quick checks. Use
larger `nboot` for more precise inference.

**Example 3: With covariate, interaction model, bootstrap inference**

```{r, error = TRUE}
persuasio4ytz2lpr(
  data   = GKB,
  y      = "voteddem_all",
  t      = "readsome",
  z      = "post",
  x      = "MZwave2",
  model  = "interaction",
  method = "bootstrap",
  level  = 0.80,
  nboot  = 1000
)
```

With covariates, method = "bootstrap"` is required. The `model = "interaction"`
specification allows the effect of `MZwave2` to differ freely between the
`z = 1` and `z = 0` subgroups.

#### Relationship to `persuasio`

```{r eval=FALSE}
# These two calls are equivalent
persuasio(
  est     = "lpr",
  varlist = c("voteddem_all", "readsome", "post"),
  data    = GKB,
  level   = 0.80,
  method  = "normal"
)

persuasio4ytz2lpr(
  data   = GKB,
  y      = "voteddem_all",
  t      = "readsome",
  z      = "post",
  method = "normal",
  level  = 0.80
)
```

### `persuasio4yz`: APR Bounds Without a Treatment Variable

When `est = "yz" in `persuasio`, persuasio4yz` is working under the hood. It
estimates bounds on the APR using only the binary outcome \code{y} and binary
instrument \code{z} and is appropriate when treatment uptake data are
unavailable or when the researcher wants to characterise the reduced-form
relationship between instrument and outcome alone.

### When to use `persuasio4yz` vs `persuasio4ytz`

| Function          | Variables needed  | Estimand          | Use when                              |
|-------------------|-------------------|-------------------|---------------------------------------|
| `persuasio4ytz`   | `y`, `t`, `z`     | APR (tighter bounds) | Treatment data available           |
| `persuasio4yz`    | `y`, `z`          | APR (wider bounds)   | No treatment data                  |

#### Examples using the GKB dataset
**Example 1: No covariates, normal inference**

```{r, error = TRUE}
persuasio4yz(
  data   = GKB,
  y      = "voteddem_all",
  z      = "post",
  method = "normal",
  level  = 0.80
)
```

Without covariates, analytic standard errors are available from both `aprlb`
and `aprub`, and the Stoye (2009) normal approximation is feasible.

**Example 2: No covariates, bootstrap inference**

```{r, error = TRUE}
persuasio4yz(
  data   = GKB,
  y      = "voteddem_all",
  z      = "post",
  method = "bootstrap",
  level  = 0.80,
  nboot  = 1000
)
```

Bootstrap inference jointly resamples the lower and upper bound estimators.
Results should align with Example 1 in large samples. The default `nboot = 50`
is only appropriate for quick checks. Use larger `nboot` for more precise
inference.

**Example 3: With covariate, interaction model, bootstrap inference**

```{r, error = TRUE}
persuasio4yz(
  data   = GKB,
  y      = "voteddem_all",
  z      = "post",
  x      = "MZwave2",
  model  = "interaction",
  method = "bootstrap",
  level  = 0.80,
  nboot  = 1000
)
```

With covariates, `method = "bootstrap"` is required. The `model = "interaction"`
specification allows heterogeneous covariate effects of `MZwave2` across
instrument groups.

#### Relationship to `persuasio`

```{r eval=FALSE}
# These two calls are equivalent
persuasio(
  est     = "yz",
  varlist = c("voteddem_all", "post"),
  data    = GKB,
  level   = 0.80,
  method  = "normal"
)

persuasio4yz(
  data   = GKB,
  y      = "voteddem_all",
  z      = "post",
  method = "normal",
  level  = 0.80
)
```
